### 2019-05-22 thesis-template/msc-thesis.tex
update

Tensorfelder werden meistens in Verbindung mit mechanischen Spanunngsverteilungen in 2D/3D-Gittern gebracht (vgl. Cauchy-Stress Tensor), haben aber auch andere praktische Bedeutungen in der Physik.
Mithilfe von globalen Beleuchtungsmodellen/-techniken wird eine neue Methode entwickelt, um Tensorfelder zu visualisieren. Als Grundlage leiten wir ein einfaches Lichtausbreitungsschema für kartesische Gitter ab, dass die Prinzipien der Ausbreitungsdämpfung und Energieerhaltung beachtet und Licht-verteilung/en für gegebene Lichtquellenrichtung/en und -positionen bis zur Konvergenz approximieren kann. Als unterliegendes Modell werden die Transmissionsprofile innerhalb diesem Gitter als Kristallfiberstruktur angenommen (vgl. Edelsteine: Katzenaugeneffekt beim Tigerauge). Die folgende Aufgabe ist es, anisotrope Fiberstrukturen mit der Orientierung und den Anisotropiemaßen des unterliegenden Tensorfelds zu modellieren. Dafür erstellen wir per Hauptachsentransformation (PCA) ein Eigensystem für jede Zelle des Tensorfelds und modellieren eine zugehörige Ellipsengleichung als (Transfer-/Wichtungs-) Transmissionsprofil. Wir messen Impulsantworten des Tensorfelds mit Delta-Pulsen an jeder Position und in jede Richtung um eine uniform gesamplete Map der globalen Lichtverteilungen zu erhalten. Diese Map wird als ein globales Energieflussfeld betrachtet. Wir lassen uns von der Technik FTLE aus dem Bereich Particle Tracing der Vector Field Visualization inspirieren, indem wir den Gradient des Flussfelds der resultierenden Lichtverteilungen analysieren und visualisieren. Ähnlich wie FTLE in Vektorfeldern Gräte/Kämme/LCS detektiert, erfasst unsere Methode dieselben Strukturen, anziehende/abstoßende/sattelnde Tensorfeldlinien die häufig Schlüsselstrukturen in der Natur repräsentieren, in Tensorfeldern. Wir führen diese Größe als Global Illumination Gradient ein, der einen FTLE-verwandten Ansatz darstellt, um LCS in Tensorfeldern durch die Analyse des globalen, durch einen Imprint (Gravur in kristallinen Fiberstrukturen) gerichteten Lichtflusses, zu visualisieren. Außerdem führen wir Deformation Ellpsoid Glyphen ein, die zusätzlich zur Darstellung der anisotropen Topologie die Richtung(Zug-/Druck) der Spannung durch Pfeile nach innen/außen darstellen. Das Verfahren wird in einer extensiven Evaluationsphase auf Plausibilität und Vergleichbarkeit geprüft.
\newpage

\chapter*{Abstract}
%\input{abstract}
Tensor Fields, are most commonly associated with stress distributions (cnf. Cauchy-Stress Tensor) in 2D/3D grids, but also have some other meanings in practice in physics. By means of global illumination techniques we develop a new method to visualize tensor fields. As a basis, we derive a simple light propagation scheme for Cartesian grids, which satisfies propagation attenuation and energy-conservation principles and is able to approximate light distributions for given light source position(s) and direction(s) until convergence. The transmission profiles within this grid are considered as crystal fiber structures as an underlying physical model (eg. gemstones: tiger's eye's cat's eye effect). The consequent task is to model anisotropic fibre structures with the orientation and anisotropy measures of the underlying tensor field in the grid.
For this, we derive an eigenbasis by PCA (principal component analysis) for every cell of the tensor field and form an ellipsoid equation as transmission (transfer/weighting) function for the transmitted light profiles. We measure impulse responses of the tensor field with Delta-Pulses as light sources at any (sampled) position and in any direction to generate a uniform sampled map of the global illumination distributions. This map is considered as a global illumination energy flow field.
We gain inspiration by vector field particle tracing's FTLE approach for analyzing the gradient of the flow field of the resulting light distributions. Similar to how the FTLE detects ridges/LCS (Lagrangian Coherent Structures) in vector fields (flow fields), our approach captures these structures likewise, attracting/repelling/saddling tensor field lines (TFL) which frequently represent key structures in nature, in tensor fields. We denote this entity a global illumination gradient which states an FTLE-related approach for visualizing LCS in tensor fields through analysis of the global and light transport  directed by imprint/engraving in crystelline fiber structures. 
We also introduce Deformation Ellipsoid Glyphs which are, additionally to visualizing anisotropic topology, pointing inward/outward for compressive/tensile stresses visualizing orientation of principal mechanical stresses.
We evaluate our approaches for plausibility and comparability within an extensive test campaign.
\newpage

@@

In this work, we wil discuss the advantages of tensor field analysis and develop a new method for tensor field visualization. At first, we will shortly state the problem setting objectives and then give a quick overview about the structure of this work.

Dieses Kapitel gibt einen Überblick über die Arbeit. Gerade der
Abschnitt zur Motivation soll allgemein verständlich geschrieben
werden. Die Einleitung sollte auch wichtige Referenzen enthalten.

@@

Tensor Field Visualization is gaining importance as a relevant tool for the analysis of fluid and solid mechanics. Tensors are found most commonly in medical, scientific and engineering applications. A tensor is a $n$-D generalization of a matrix and appears e.g. as Jacobi-Matrix in Flow Fields or as stress tensor in solid continuum mechanics. Basically, tensors describe stretchings, rotations and volume changes in both solid and fluid material. While most techniques focus on symmetric tensor field visualization (like Glyphs and Tensor Field Lines), some other recent works adress the problem of asymmetric tensor field visualization. The issue with asymmetric or antisymmetric tensors is that they do not always yield real eigenvalues, which is needed to determine an eigenbasis to set up glyphs, hyperstreamlines or tensor field lines. \cite{pang&zheng} et.al. proposed the concept of dual eigenvectors and \cite{zhang} extended it to pseudo-eigenvectors and introduced eigenvector and eigenvalue manifold to visualize eigenvectors in the complex domain. \\

\begin{figure}[h]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/hyperstreamlines.png}
  }
  \caption{Hyperstreamlines in Tensor Field Visualization}
  \label{Skalierung}
\end{figure}
Worum geht es? Beispiel(e)! Illustrationen sind hier meist sinnvoll
zum Verständnis. Warum ist das Thema wichtig? In welchem Kontext?

\section{Objectives}
The problem setting encomprises the visualization of tensor fields, i.e. mostly 2D/3D-grids consisting of mainly second-order tensors (cnf. Cauchy-Stress Tensor). The main purpose of visualization in general is to reduce or encode some set of numbers in some readable, explorable and more intuitive representation. This implies e.g. clustering, structuring or projection techniques. The objective of this work is to develop a simple light propagation scheme to generate a tensor-induced flow map. The tensors are decomposed via singular value decomposition to obtain a real eigenbasis as transmission profile. Another aim is to implement a vector-field visualization borrowed FTLE-related approach in tensor field visualization (both symmetric and asymmetric) for detecting ridges/LCS (Lagrangian Coherent Structures). These structures can be identified as repelling, attracting and saddling (forming a saddle) local features, which frequently represent key structures in nature. LCS usually separate regions of different flow behavior, but in our case they separate regions/domains of differing stress distributions. The map responds most where tensor field lines / hyperstreamlines converge or diverge, whereas one is the same as the other because of bidirectional stress directions. The real part of the eigenvalues can be exploited for sign change (contractions/tensions) which leads to the concept of deformation glyphs w. attached arrows pointing inward/outward for the two types of normal stresses. Practically, in its eigenbasis there are no shear stresses existent for the symmetric stress tensor in equilibrium, just the principal normal stresses. Thus, we can even define the full set of stresses with the deformation glyphs which are derived from the singular value decomposition and signed with the real part of the eigenvalues (in corresponding absolute order).

@@

For the rest of this work, we will give a quick overview about the structure and outline. Next, in chap. \ref{chap:grundlagen}, we will capture important fundamentals like Tensors and PCA in the context of singular value decomposition. Also we will discuss the most relevant work in the domain of tensor field visualization. In chap. \ref{chap:main} we will propose the main contribution of this work and document the experimental evaluation in chap. \ref{chap:eval}. The Summary, Conclusion and Outlook will be given in chap. \ref{chap:concl}.\\

Dieser Abschnitt wird meist recht kurz gehalten und beschreibt im
Prinzip nur den Aufbau des Rests der Arbeit. Zum Beispiel: In Kapitel
\ref{chap:grundlagen} geben wir einen Überblick über die  Grundlagen
zu der Arbeit sowie über verwandte Arbeiten. In Kapitel
\ref{chap:main} stellen wir dann \ldots vor. \ldots etc.

@@

Die ersten paar Abschnitte in diesem Kapitel führen in die Grundlagen
zur Arbeit ein. Das können beispielsweise Grundlagen zu Netzwerken
oder zur Informationsextraktion sein.

\section{Tensors}
\label{sec:tensors}
Now what is a Tensor? A tensor is a $o$-dimensional generalization of a matrix as follows:
\begin{center}
 \begin{tabular}{c|c|c|c|c|c|c}
\textbf{order} & 0 & 1 & 2 & 3 & ... & o \\ 
\hline 
\textbf{shape} & scalar & vector & matrix & 3D-matrix & ... & oD-matrix\\ 
\end{tabular}\vskip 5pt
Table 1: Tensor Shapes
 \end{center}

In addition to that, a tensor has as many indices as its order $o$ and their run length is as long as its embedded dimension $n$, which is equal to the rank of the tensor for full rank. Tensors follow certain transformation rules which are defined for covariant/contravariant tensors. They are most commonly obtained and defined from physical transformation equations. Hence, any $o$D number array could be a tensor, but the definition holds only if the transformation rules apply to these $o$D number arrays (in a sense of physical interpretation), which is mostly true for component-indexed $o$D number entities following matrix multiplication and transformation rules found in physics and math. In fact, tensors themselves are interpreted as a transformation which reflects an incoming direction vector (measuring e.g. the stress in that direction) to an outgoing stress vector.
 To be a bit more concrete, tensors represent stress distributions in solids and fluids describing strain and flow features as these entities aren't simple numbers or vectors. These numbers just come up as entities or quantities in physics. That is, at each point in space (the grid) there is a whole distribution of e.g. stresses (elastic and viscous) which needs to be characterized and represented.
 \\
 \emph{Cauchy stress tensor}
 \begin{figure}[h]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/stress_tensor.png}
  }
  \caption{Cauchy-Stress Tensor}
  \label{cauchy}
\end{figure}

The Cauchy stress tensor, which is propably the classical example of a tensor, is depicted in Fig. \ref{cauchy} and consists of 3 stress vectors $\mathbf{T(e_i)}$ arranged in row-major order. It appears in similar form as viscous stress tensor in fluid mechanics:
\begin{align*}
\boldsymbol{\sigma}=
\begin{bmatrix}
\sigma _{11} & \sigma _{12} & \sigma _{13} \\
\sigma _{21} & \sigma _{22} & \sigma _{23} \\
\sigma _{31} & \sigma _{32} & \sigma _{33} 
\end{bmatrix}
\end{align*}
These 3 stress vectors represent the orientation and magnitude of total resulting stress at plane ${x,y,z}$ in direction ${x,y,z}$ (9 individual numbers). Thus, the stress tensor poses a full valid represenation of the stress distribution at any point in space. The stress tensor itself can be interpreted as a transformation which maps an incoming normal direction vector $\mathbf{n}$ a resulting stress vector $\mathbf{T(n)} = \mathbf{n}\cdot\boldsymbol{\sigma}$. The stresses in normal directions (diagonal elements) are directly related to the pressure, which is always isotropic and hence non-directionally dependent. It is possible to find 3 principal stress directions in 3D, for which there are no shear stresses existent in equilibrium, by PCA (see sect. \ref{sec:pca} below). These form the principal directions of deformation at a certain location and are sufficient to describe the resulting transformation/deformation behaviour. The resulting system of principal directions and absolute stresses is often called an eigensystem or eigenframe. So to say, an eigensystem consjstent of the principal directions/axes spans a principal ellipsoid depicted in Fig. \ref{HAT} in sect. \ref{sec:pca}.

\section{Principal Component Analysis}
\label{sec:pca}
Note that we denote a matrix as a second order tensor interchangeably for the sake of simplicity in the following. For any state of stress in equilibrium it is possible to find $n$ (dimensionality) independent orthogonal directions with no shear stresses existent. Along these directions the principal stresses, which are normal (tensile/compressive) stresses, are exerted. The anisotropy can be represented by ellipsoid glyphs.
\begin{figure}[h]
  \centering
 {
    \includegraphics[width=0.5\textwidth]
    {img/HAT.png}
  }
  \caption{Principal Component Analysis}
  \label{HAT}
\end{figure}

Principal Component Analysis is an algorithm which can capture the principal components (directions) in stochastic data. But it works similarly on transformations to find the principal directions on rotation and scaling. Now, principal component analysis can be done by eigenvalue or singular vector decomposition, whereas the former yields complex eigenvalues for asymmetric matrices. For (anti-)symmetric (normal) matrices, they are connected through the following relation: $s_i = \sqrt{|\lambda_i|}$. Mathematically speaking, a decomposition is somehow decomposing a transformation (which follows certain transformation rules) into parts of subsequent transformations consistent of rotation and scaling ($\mathbf{A} = \mathbf{R}\mathbf{S}\mathbf{R}^*$). We choose to use singular value decomposition ($\mathbf{A} = \mathbf{U} \mathbf{\Sigma} \mathbf{V}^*$) to be able to set up real eigenvalues for any kind of matrix (singular values are always real and positive). Also, the singular vectors of the $\mathbf{U}$-matrix represent the half-axes of the principal ellipsoid (eigenframe) shown in Fig. \ref{HAT} for any kind of matrix as suggested by \cite{moler}. In addition, we also calculate the eigenvalues to exploit the sign of the real part of the eigenvalues ordered decreasingly by absolute value (corresponding to singular value order). The sign corresponds to tensile ($+$) and compressive ($-$) stress, which allows us to visualize deformation glyphs with arrows pointing inwards for compressive and outwards for tensile stresses. We do this decomposition analysis once for each matrix and then store the precomputed results in a grid. The principal ellipsoid is also used as a transmission profile for the propagation scheme in symbolic form in a subequent step.
\vskip 5pt
\section{Related Work}
\label{sec:related}

@@

\textbf{Symmetric Tensor Field Visualization}\\
\cite{pang&zheng} et.al. proposed a texture-based visualization approach called HyperLIC, which extends the concept of LIC to symmetric tensor fields by using an anisotropic 2D-filter kernel oriented along the major/minor eigenvectors. The notion of tensor field topology and the concept of Hyperstreamlines was first introduced by \cite{delmarcelle&hesselink}. The placement of tensor field lines as hyperstreamlines was recently improved by \cite{spencer} et.al. for glyph packing. A good overview of seeding strategies is given by \cite{mcloughlin}. \cite{feng} et. al. used Voronoi Tesselation for placing the glyphs. Superquadric glyphs have been proposed by \cite{schultz&kindlmann} for general (non-positive definite) tensors. \cite{deleeuw&vanwijk} visualize the partial derivative gradient, the Jacobian of the tensor field to sense the local properties of the field. 
\\
\textbf{Asymmetric Tensor Field Visualization}\\
\cite{pang&zheng} et.al. proposed the concept of dual eigenvectors and \cite{zhang} extended it to pseudo-eigenvectors and introduced the eigenvector/eigenvalue manifolds to visualize eigenvectors in the complex domain. \cite{laramee} et.al. focused on the efficient implementation and visualization of these structures and provided an interactive visualization system for asymmetric tensors applicable in fluid and solid dynamics. In this work, we will use the singular vectors yielded from singular value decomposition to avoid complex values yielded from eigenvalue decomposition. This will allow us to set up glyphs representing the major/minor axis of the underlying transformation for every kind of (asymmetric/symmetric) tensor. We will then use global-illumination techniques to take it one step further and to generate a flow map, permitting us to compute an FTLE-like field on the tensor fields. The concept of Tensor Magnitude has been introduced by \cite{laramee2} for means of physical interpretation. They also proposed an efficient glyph and hyperstreamline hybrid approach, which made dynamic interaction in real-time in 2D tensor fields feasible. \cite{palacios} extracted isosurfaces of tensor magnitude, mode and isotropy index.

@@

\begin{figure}[h]
  \centering
 {
    \includegraphics[width=1.0\textwidth]
    {img/steps_gallery_format_arrow_neg.png}
  }
  \caption{Impulse Response for Isotropic Stimulus (Few Iterations: 0,1,2,3,7,11)}
  \label{iter}
\end{figure}\vskip 3pt

>>>>

Tensorfelder werden meistens in Verbindung mit mechanischen Spanunngsverteilungen in 2D/3D-Gittern gebracht (vgl. Cauchy-Stress Tensor), haben aber auch andere praktische Bedeutungen in der Physik.
Mithilfe von globalen Beleuchtungsmodellen/-techniken wird eine neue Methode entwickelt, um Tensorfelder zu visualisieren, die aber zusätzlich auch verwendet werden kann um die Lichtausbreitung in einer topologischen Szene zu beschreiben. Als Grundlage leiten wir ein einfaches Lichtausbreitungsschema für kartesische Gitter ab, dass die Prinzipien der Ausbreitungsdämpfung und Energieerhaltung beachtet und Licht-verteilung/en für gegebene Lichtquellenrichtung/en und -positionen bis zur Konvergenz approximieren kann. Als unterliegendes Modell werden die Transmissionsprofile innerhalb diesem Gitter als Kristallfiberstruktur angenommen (vgl. Edelsteine: Katzenaugeneffekt beim Tigerauge). Die folgende Aufgabe ist es, anisotrope Fiberstrukturen mit der Orientierung und den Anisotropiemaßen des unterliegenden Tensorfelds zu modellieren. Dafür erstellen wir per Hauptachsentransformation (PCA) ein Eigensystem für jede Zelle des Tensorfelds und modellieren eine zugehörige Ellipsengleichung als (Transfer-/Wichtungs-) Transmissionsprofil. Folglich kann die resultierende Lichtverteilung als 2D Skalarfeld oder Polarplots für Richtungsinformation visualisiert/repräsentiert werden. Wir messen Impulsantworten des Tensorfelds mit Delta-Pulsen an jeder Position und in jede Richtung um eine uniform gesamplete Map der globalen Lichtverteilungen zu erhalten. Diese Map wird als ein globales Energieflussfeld betrachtet. Wir lassen uns von der Technik FTLE aus dem Bereich Particle Tracing der Vector Field Visualization inspirieren, indem wir den Gradient des Flussfelds der resultierenden Lichtverteilungen analysieren und visualisieren. Ähnlich wie FTLE in Vektorfeldern Gräte/Kämme/LCS detektiert, erfasst unsere Methode dieselben Strukturen, anziehende/abstoßende/sattelnde Tensorfeldlinien die häufig Schlüsselstrukturen in der Natur repräsentieren, in Tensorfeldern. Wir führen diese Größe als Global Illumination Gradient ein, der einen FTLE-verwandten Ansatz darstellt, um LCS in Tensorfeldern durch die Analyse des globalen, durch einen Imprint (Gravur in kristallinen Fiberstrukturen) gerichteten Lichtflusses, zu visualisieren. Außerdem führen wir Deformation Ellpsoid Glyphen ein, die zusätzlich zur Darstellung der anisotropen Topologie die Richtung(Zug-/Druck) der Spannung durch Pfeile nach innen/außen darstellen. Das Verfahren wird in einer extensiven Evaluationsphase auf Plausibilität und Vergleichbarkeit geprüft.
\newpage

\chapter*{Abstract}
%\input{abstract}
Tensor Fields, are most commonly associated with stress distributions (cnf. Cauchy-Stress Tensor) in 2D/3D grids, but also have some other meanings in practice in physics. By means of global illumination techniques, we develop a new method to visualize tensor fields, which is also capable of visualizing the light transport (propagation) in a topologically defined scene geometry. As a basis, we derive a simple light propagation scheme for Cartesian grids, which satisfies propagation attenuation and energy-conservation principles and is able to approximate light distributions for given light source position(s) and direction(s) until convergence. The transmission profiles within this grid are considered as crystal fiber structures as an underlying physical model (eg. gemstones: tiger's eye's cat's eye effect). The consequent task is to model anisotropic fibre structures with the orientation and anisotropy measures of the underlying tensor field in the grid.
For this, we derive an eigenbasis by PCA (principal component analysis) for every cell of the tensor field and form an ellipsoid equation as transmission (transfer/weighting) function for the transmitted light profiles. Hence, the resulting light distribution can be visualized via 2D scalar field or polar plots for directional information. Additionally, We measure impulse responses of the tensor field with Delta-Pulses as light sources at any (sampled) position and in any direction to generate a uniform sampled map of the global illumination distributions. This map is considered as a global illumination energy flow field.
We gain inspiration by vector field particle tracing's FTLE approach for analyzing the gradient of the flow field of the resulting light distributions. Similar to how the FTLE detects ridges/LCS (Lagrangian Coherent Structures) in vector fields (flow fields), our approach captures these structures likewise, attracting/repelling/saddling tensor field lines (TFL) which frequently represent key structures in nature, in tensor fields. We denote this entity a global illumination gradient which states an FTLE-related approach for visualizing LCS in tensor fields through analysis of the global and light transport  directed by imprint/engraving in crystelline fiber structures. 
We also introduce Deformation Ellipsoid Glyphs which are, additionally to visualizing anisotropic topology, pointing inward/outward for compressive/tensile stresses visualizing orientation of principal mechanical stresses.
We evaluate our approaches for plausibility and comparability within an extensive test campaign.
\newpage

@@

In this work, we wil discuss the advantages of tensor field analysis and develop a new method for tensor field and a coherent light transport visualization. At first, we will shortly state the problem setting objectives and then give a quick overview about the structure of this work.

%Dieses Kapitel gibt einen Überblick über die Arbeit. Gerade der
%Abschnitt zur Motivation soll allgemein verständlich geschrieben
%werden. Die Einleitung sollte auch wichtige Referenzen enthalten.

@@

Tensor Field Visualization is gaining importance as a relevant tool for the analysis of fluid and solid mechanics. Tensors are found most commonly in medical, scientific and engineering applications. A tensor is a $n$-D generalization of a matrix and appears e.g. as Jacobi-Matrix in Vector (Flow) Field Visualization or as stress tensor in solid continuum mechanics. Basically, tensors describe compressions, tensions, rotations and volume changes in both solid and fluid material. While most techniques focus on symmetric tensor field visualization (like Glyphs and Tensor Field Lines), some other recent works adress the problem of asymmetric tensor field visualization. The issue with asymmetric or antisymmetric tensors is that they do not always yield real eigenvalues, which is needed to determine an eigenbasis to set up glyphs, hyperstreamlines or tensor field lines. \cite{pang&zheng} et.al. proposed the concept of dual eigenvectors and \cite{zhang} extended it to pseudo-eigenvectors and introduced eigenvector and eigenvalue manifold to visualize eigenvectors in the complex domain. In this work, we will use the $\mathbf{U}$-matrix obtained from singular value decomposition (SVD) to obtain the half-axes of the PCA principal ellipsoid as suggested by \cite{moler}.\\

\begin{figure}[h]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/DIFFERENT_TYPES_OF_STRESS.png}
  }
  \caption{Different types of stresses}
  \label{Skalierung}
\end{figure}

Worum geht es? Beispiel(e)! Illustrationen sind hier meist sinnvoll
zum Verständnis. Warum ist das Thema wichtig? In welchem Kontext?

\section{Objectives}
\begin{figure}[h]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/hyperstreamlines.png}
  }
  \caption{Hyperstreamlines in Tensor Field Visualization}
  \label{Skalierung}
\end{figure}
The problem setting encomprises the visualization of tensor fields, i.e. mostly 2D/3D-grids consisting of mainly second-order tensors (cnf. Cauchy-Stress Tensor) with the additional possibility to model light transport transport in a geometrical scene. The main purpose of visualization in general is to reduce or encode some set of numbers in some readable, explorable and more intuitive representation. This implies e.g.clustering, structuring or projection techniques. The objective of this work is to develop a simple light propagation scheme to generate a tensor-induced flow map. The tensors are decomposed via singular value decomposition to obtain a real eigenbasis as transmission profile. Another aim is to implement a vector-field visualization borrowed FTLE-related approach in tensor field visualization (both symmetric and asymmetric) for detecting ridges/LCS (Lagrangian Coherent Structures). These structures can be classified (identified) as repelling, attracting and saddling (forming a saddle) local features, which frequently represent key structures in nature. LCS usually separate regions of different flow behavior, but in our case they separate regions/domains of differing stress distributions. The map responds most where tensor field lines / hyperstreamlines converge or diverge, whereas one is the same as the other because of bidirectional stress directions. The real part of the eigenvalues can be exploited for sign change (contractions/tensions) which leads to the concept of deformation glyphs w. attached arrows pointing inward/outward for the two types of normal stresses. Practically, in its eigenbasis there are no shear stresses existent for the symmetric stress tensor in equilibrium, just the principal normal stresses. Thus, we can even define the full set of stresses with the deformation glyphs which are derived from the singular value decomposition and signed with the real part of the eigenvalues (in corresponding absolute order).

@@

For the rest of this work, we will give a quick overview on the structure and outline. Next, in chap. \ref{chap:grundlagen}, we will capture important fundamentals like Tensors and PCA in the context of singular value decomposition (generalization of eigenvalue decomposition). Also we will discuss the most relevant work in the domain of tensor field visualization. In chap. \ref{chap:main} we will propose the main contributions of this work and document the experimental evaluation in chap. \ref{chap:eval}. A compact Summary, Conclusion and Outlook will be given in chap. \ref{chap:concl}.\\

%Dieser Abschnitt wird meist recht kurz gehalten und beschreibt im
%Prinzip nur den Aufbau des Rests der Arbeit. Zum Beispiel: In Kapitel
%\ref{chap:grundlagen} geben wir einen Überblick über die  Grundlagen
%zu der Arbeit sowie über verwandte Arbeiten. In Kapitel
%\ref{chap:main} stellen wir dann \ldots vor. \ldots etc.

@@

%Die ersten paar Abschnitte in diesem Kapitel führen in die Grundlagen
%zur Arbeit ein. Das können beispielsweise Grundlagen zu Netzwerken
%oder zur Informationsextraktion sein.

\section{Tensors}
\label{sec:tensors}
Now what is a Tensor? A tensor is a $o$-dimensional generalization of a matrix as follows:
\begin{center}
 \begin{tabular}{c|c|c|c|c|c|c}
\textbf{order} & 0 & 1 & 2 & 3 & ... & o \\ 
\hline 
\textbf{shape} & scalar & vector & matrix & "3D-matrix" & ... & oD-matrix\\ 
\end{tabular}\vskip 5pt
Table 1: Tensor Shapes
 \end{center}

In addition to that, a tensor has as many indices as its order $o$ and their run length is as long as its embedded dimension $n$, which is equal to the rank of the tensor for full rank (or in other words the number of elements n for a $nxn$-matrix). Tensors follow certain transformation rules which are defined for covariant/contravariant tensors. They are most commonly obtained and defined from physical transformation equations. Hence, any $o$-D number array could be a tensor, but the definition holds only if the transformation rules apply to these $o$-D number arrays (in a sense of physical interpretation), which is mostly true for component-indexed $o$D number entities following matrix multiplication and transformation rules found in physics and math. In fact, tensors themselves are interpreted as a transformation which reflects an incoming direction vector (measuring e.g. the stress in that direction) to an outgoing stress vector.
 To be a bit more concrete, tensors represent stress distributions in solids and fluids describing strain and flow features as these entities aren't simple numbers or vectors. They occur in form of a tensor fields, and thus as entities or quantities in physics. That is, at each point in space (the grid) there is a whole distribution of e.g. stresses (elastic and viscous) which needs to be characterized and represented by a single tensor.
 \\
 \emph{Cauchy stress tensor}
 \begin{figure}[h]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/stress_tensor.png}
  }
  \caption{Cauchy-Stress Tensor}
  \label{cauchy}
\end{figure}

The Cauchy stress tensor, which is propably the classical example of a tensor, is depicted in Fig. \ref{cauchy} and consists of 3 stress vectors $\mathbf{T(e_i)}$ arranged in row-major order. It appears in similar form as viscous stress tensor in fluid mechanics:
\begin{align*}
\boldsymbol{\sigma_{ij}}
 = \left[{\begin{matrix} \mathbf{T}^{(\mathbf{e}_1)} \\
\boldsymbol{T}^{(\mathbf{e}_2)} \\
\boldsymbol{T}^{(\mathbf{e}_3)} \\
\end{matrix}}\right] =
\left[{\begin{matrix}
\sigma _{xx} & \sigma _{xy} & \sigma _{xz} \\
\sigma _{yx} & \sigma _{yy} & \sigma _{yz} \\
\sigma _{zx} & \sigma _{zy} & \sigma _{zz} \\
\end{matrix}}\right]  = \left[{\begin{matrix}
\sigma _x & \tau _{xy} & \tau _{xz} \\
\tau _{yx} & \sigma _y & \tau _{yz} \\
\tau _{zx} & \tau _{zy} & \sigma _z \\
\end{matrix}}\right] =
\begin{bmatrix}
\sigma _{11} & \sigma _{12} & \sigma _{13} \\
\sigma _{21} & \sigma _{22} & \sigma _{23} \\
\sigma _{31} & \sigma _{32} & \sigma _{33} 
\end{bmatrix}
\end{align*}
These 3 stress vectors represent the orientation and magnitude of total resulting stress at plane ${x,y,z}$ in direction ${x,y,z}$ (9 individual numbers). Thus, the stress tensor poses a full valid represenation of the stress distribution at any point in space. The stress tensor itself can be interpreted as a transformation which maps an incoming normal direction vector $\mathbf{n}$ a resulting stress vector $\mathbf{T^{(n)}} = \mathbf{n}\cdot\boldsymbol{\sigma}$. The stresses in normal directions (diagonal elements) are directly related to the pressure, which is always isotropic and hence non-directionally dependent. It is possible to find 3 principal stress directions in 3D through PCA (see sect. \ref{sec:pca} below), for which there are no shear stresses existent in equilibrium. These form the principal directions of deformation at a certain location and are sufficient to describe the resulting transformation/deformation behaviour. The resulting system of principal directions and absolute stresses is often called an "eigensystem" or "eigenframe". So to say, an eigensystem, consistent of the principal directions/axes, spans a principal ellipsoid depicted in Fig. \ref{HAT} in sect. \ref{sec:pca}.

\section{Principal Component Analysis}
\label{sec:pca}
Note that we denote a matrix as a second order tensor interchangeably for the sake of simplicity in the following. For any state of stress in equilibrium it is possible to find $n$ (dimensionality) independent orthogonal directions with no shear stresses existent. Along these directions the principal stresses, which are normal (tensile/compressive) stresses, are exerted. The anisotropy can be represented by e.g. ellipsoid glyphs.

\begin{figure}[h]
%\centering
  \begin{minipage}{0.5\textwidth}
    \includegraphics[width=\textwidth]{img/HAT.png}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.5\textwidth}
    \includegraphics[width=\textwidth]{img/glyphs.png}
    \label{b)}
  \end{minipage}
\caption{a) Principal Component Analysis, b) Ellipsoid Glyphs w. anisotropy coefficient (l:linear, p:planar, s:spherical)}
\label{gradient}
\end{figure}

Principal Component Analysis is an algorithm which can capture the principal components (directions) in stochastic data. But it works similarly on transformations to capture the principal directions from transformations like rotation and scaling. Now, principal component analysis can be done by eigenvalue or singular vector decomposition, whereas the former yields complex eigenvalues for asymmetric matrices. For (anti-)symmetric (normal) matrices, they are connected through the following relation: $s_i = \sqrt{|\lambda_i|}$ (cnf. \cite{kieburg}). Mathematically speaking, a decomposition is somehow decomposing a transformation (which follows certain transformation rules) into parts of subsequent transformations consistent of rotation and scaling ($\mathbf{A} = \mathbf{R}\mathbf{S}\mathbf{R}^*$). We choose to use singular value decomposition ($\mathbf{A} = \mathbf{U} \mathbf{\Sigma} \mathbf{V}^*$) to be able to set up real eigenvalues for any kind of matrix (singular values are always real and positive). Also, the singular vectors of the $\mathbf{U}$-matrix represent the half-axes of the principal ellipsoid (eigenframe) shown in Fig. \ref{HAT} for any kind of matrix as pointed out by \cite{moler}. In addition, we also calculate the eigenvalues to exploit the sign of the real part of the eigenvalues (\cite{mroz}) ordered decreasingly by absolute value (corresponding to singular value order). The sign corresponds to tensile ($+$) and compressive ($-$) stress, which would allow us to visualize deformation glyphs with arrows pointing inwards for compressive and outwards for tensile stresses. We do this decomposition analysis once for each matrix and then store the precomputed results in a grid. The principal ellipsoid is also used as a transmission profile for the propagation scheme in symbolic form in a subequent step. In this work, we will use the singular vectors yielded from singular value decomposition to avoid complex values yielded from eigenvalue decomposition. This will allow us to set up glyphs representing the major/minor axis of the underlying transformation for every kind of (asymmetric/symmetric) tensor. We will then use global-illumination techniques on top to generate a flow map, which permits us to compute an FTLE-like field on the tensor fields.
\vskip 5pt
\section{Related Work}
\label{sec:related}

@@

\textbf{Symmetric Tensor Field Visualization}\\
\cite{pang&zheng} et.al. proposed a texture-based visualization approach called HyperLIC, which extends the concept of LIC to symmetric tensor fields by using an anisotropic 2D-filter kernel oriented along the major/minor eigenvectors. The concept of following the major/minor eigenvector along tensor field lines was invented by \cite{vilanova}. Tensorlines introduce a kind of artificial inertia (running average), making them resistent to noise. The notion of tensor field topology and the concept of Hyperstreamlines was first introduced by \cite{delmarcelle&hesselink}. The placement of tensor field lines as hyperstreamlines was recently improved by \cite{spencer} et.al. for glyph packing. A good overview of seeding strategies is given by \cite{mcloughlin}. \cite{feng} et. al. used Voronoi Tesselation for placing the glyphs. Superquadric glyphs have been proposed by \cite{schultz&kindlmann} for general (non-positive definite) tensors. \cite{deleeuw&vanwijk} visualize the partial derivative gradient, the Jacobian of the tensor field to sense the local properties of the field. A diffusion tensor can also be visualized by box, ellipsoid \cite{pierpaoli}, composite \cite{worth} or superquadric \cite{kindlmann} glyphs. There is also a set of scalar measures available in symmetric tensor field visualization such as: mean diffusivity, fractional anisotropy \cite{basser} and anisotropy coefficients \cite{westin}.
\\
\textbf{Asymmetric Tensor Field Visualization}\\
\cite{pang&zheng} et.al. proposed the concept of dual eigenvectors and \cite{zhang} extended it to pseudo-eigenvectors and introduced the eigenvector/eigenvalue manifolds to visualize eigenvectors in the complex domain. \cite{laramee} et.al. focused on the efficient implementation and visualization of these structures and provided an interactive visualization system for asymmetric tensors applicable in fluid and solid dynamics. The concept of Tensor Magnitude has been introduced by \cite{laramee2} for means of physical interpretation. They also proposed an efficient glyph and hyperstreamline hybrid approach, which made dynamic interaction in real-time in 2D tensor fields feasible. \cite{palacios} extracted isosurfaces of tensor magnitude, mode and isotropy index.

@@

\begin{figure}[h]
  \centering
 {
    \includegraphics[width=1.0\textwidth]
    {img/steps_gallery_format_arrow_alt.png}
  }
  \caption{Impulse Response for Isotropic Stimulus (Few Iterations: 0,1,2,3,7,11)}
  \label{iter}
\end{figure}\vskip 3pt
